Using VariantSpark to learn population structure from genomic profiles

### 2015-10-26 doc/publications/Recomb15/recomb15.tex
more touchups

\FloatBarrier
\clearpage
\marginpar{TODO report bd's accuracy measure as the main measure (add ARI as the additional information linking to our
  paper)}
\subsubsection{Logistic regression and \kMeans{} capture the population better than deep learning in a three population test}
In this section we investigate whether the multivariate approach, deep learning, is able to achieve a better result than
RF by again comparing it against \kMeans{} and logistic regression.  We compare against the reported accuracy of the
deep learning approach using the {\sc ADAM} framework~\cite{Massie2013} as reported on their webpage~\cite{Ferguson}.
As the reported results were performed on only three of the 14 populations, we hence restrict our comparison also to ASW
(Americans of African Ancestry in SW USA), CHB (Han Chinese in Beijing, China), and GBR (British in England and
Scotland).

The mean ARI for our \kMeans clustering models is \(0.960 \pm 0.004 \). Although slightly lower, this value is similar to the deep-learning approach
where they achieved an ARI of 0.976 (where one ASW and one GBR individual is misclassified as CHB). Our supervised logistic regression approach,
however, improves upon this result.  With this approach, we achieve an accuracy of 100\%. This is the average across
ten folds.

To reiterate, these results are from the three populations, ASW, CHB and GBR, that belong to the super-populations, AFR,
EAS and EUR, respectively.  As we mention in the introduction, we observed a perfect clustering result when we previously clustered
the these three super-populations (rather than one population from each), with an ARI of 1.0. However, when
we introduce the fourth super-population into the mix, AMR, the ARI drops to 0.84.  Given the perfect result when building
the logistic regression model on this small subset of data, we expand our dataset to include every population from the
1000 Genomes Project phase1 dataset.

\marginpar{TODO[say that this may be a too easy dataset to benchmark the performance differences between kmeans and RF
  (hence the full dataset needed)]}

\begin{table}
\caption{Cross-validated results, mean accuracy (and standard error) on 3 populations.}
\begin{center}
\renewcommand{\arraystretch}{1.4}
\setlength\tabcolsep{3pt}
\begin{tabular}{lcccc}
\hline\noalign{\smallskip}
method  & ARI & error \\
LR  & X (Y) & -- \\
\kMeans & 0.960 (0.004) & -- \\
deep learning & 0.976 & 0.008 \\
\noalign{\smallskip}
\hline
\end{tabular}
\end{center}
\end{table}

@@

\marginpar{TODO [Discuss supervised vs unsupervised]} \marginpar{TODO can you drill down to the population clusters and
  see if and where the best performing method still makes mistakes - this will help in the conclusion to argue that
  actually the labels are wrong rather than the method not being able to capture the structure (e.g. say if AMR are
  mixed in with GBR TSI)}

>>>>

\FloatBarrier
\clearpage
\marginpar{TODO report bd's accuracy measure as the main measure (add ARI as the additional information linking to our
  paper) Aidan:See email}
\subsubsection{Logistic regression and \kMeans{} capture the population better than deep learning in a three population test}
In this section we investigate whether the multivariate approach, deep learning, is able to achieve a better result than
RF by again comparing it against \kMeans{} and logistic regression.  We compare against the reported accuracy of the
deep learning approach using the {\sc ADAM} framework~\cite{Massie2013} as reported on their webpage~\cite{Ferguson}.
As the reported results were performed on only three of the 14 populations, we hence restrict our comparison also to ASW
(Americans of African Ancestry in SW USA), CHB (Han Chinese in Beijing, China), and GBR (British in England and
Scotland).

The mean ARI for our \kMeans clustering models is \(0.960 \pm 0.004 \). Although slightly lower, this value is similar to the deep-learning approach
where they achieved an ARI of 0.976 (where one ASW and one GBR individual is misclassified as CHB). Our supervised logistic regression approach,
however, improves upon this result.  With this approach, we achieve an ARI of 1 across the ten folds.

To reiterate, these results are from the three populations, ASW, CHB and GBR, that belong to the super-populations, AFR,
EAS and EUR, respectively.  As we mention in the introduction, we observed a perfect clustering result when we previously clustered
the these three super-populations (rather than one population from each), with an ARI of 1.0. However, when
we introduce the fourth super-population into the mix, AMR, the ARI drops to 0.84.  Given the perfect result when building
the logistic regression model on this small subset of data, and the similarity with the other models,
we expand our dataset to include every population from the 1000 Genomes Project phase1 dataset.

\marginpar{TODO[say that this may be a too easy dataset to benchmark the performance differences between kmeans and RF
  (hence the full dataset needed) Aidan: done(probably)]}

\begin{table}
\caption{Cross-validated results, mean accuracy (and standard error) on 3 populations.}
\begin{center}
\renewcommand{\arraystretch}{1.4}
\setlength\tabcolsep{3pt}
\begin{tabular}{lcccc}
\hline\noalign{\smallskip}
method  & ARI & error \\
LR  & 1.000 (0.000) & 0.000 \\
\kMeans & 0.960 (0.004) & -- \\
deep learning & 0.976 & 0.008 \\
\noalign{\smallskip}
\hline
\end{tabular}
\end{center}
\end{table}

@@

\marginpar{TODO [Discuss supervised vs unsupervised] Aidan: wrote in previous section (where I mention initialisation)} \marginpar{TODO can you drill down to the population clusters and
  see if and where the best performing method still makes mistakes - this will help in the conclusion to argue that
  actually the labels are wrong rather than the method not being able to capture the structure (e.g. say if AMR are
  mixed in with GBR TSI) Aidan: not in time :( }
